Informative Oceanographic Sampling using Excursion Probabilities for Multivariate Random Fields

### 2019-09-30 main_new.tex
Trygve updates

% == Trygve Test
\usepackage{color, colortbl}
\definecolor{Gray}{gray}{0.9}
\newcommand{\kcomment}[1]{{\color{red}{\{KR: #1\}}}}
\newcommand{\tofcomment}[1]{{\color{green}{\{TF: #1\}}}}
\newcommand{\kc}{\kcomment}
% put your definitions there:
\startlocaldefs
\endlocaldefs

@@

\begin{aug}
\author{\fnms{Trygve Olav} \snm{Fossum}\thanksref{a,b}, \corref{} \ead[label=e1]{trygve.o.fossum@ntnu.no}}
\author{\fnms{Jo} \snm{Eidsvik}\thanksref{c}, \ead[label=e2]{jo.eidsvik@math.ntnu.no}}
\author{\fnms{David} \snm{Ginsbourger}\thanksref{d}, \ead[label=e3]{david.ginsbourger@idiap.ch}}
\and
\author{\fnms{Kanna} \snm{Rajan}\thanksref{b,e,f}. \ead[label=e4]{kanna.rajan@ntnu.no}}

\address{Department of Marine Technology\\ Trondheim, Norway\printead{e1}}
\affiliation[a]{Department of Marine Technology, The Norwegian University of Science and Technology (NTNU), Trondheim, Norway.} 
\affiliation[b]{Centre for Autonomous Marine Operations and Systems (AMOS).}

\address{Department of Mathematical Sciences \printead{e2}}
\affiliation[c]{Department of Mathematical Sciences, NTNU.}

\address{Idiap \printead{e3}}
\affiliation[d]{Uncertainty Quantification and Optimal Design group, Idiap Research Institute, Martigny, Switzerland.}

\address{Department of Engineering Cybernetics,\printead{e4}}
\affiliation[e]{Department of Engineering Cybernetics, NTNU}
\affiliation[f]{Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}

@@

The goal of the information-based methods that we highlight in the current paper, is to do efficient environmental sampling and overcome common challenges, namely:

\begin{itemize}
\item Full numerical ocean models cannot provide accurate results online if run on robotic sensing
  platforms, as onboard computers cannot deliver the computational power required,
  hence statistical proxy models of the environment must be used instead.
\item Decisions on where to sample happen sequentially, i.e. recently
  acquired information can be assimilated into and used to improve
  statistical models and reduce uncertainty associated with where to
  make measurements.
\item As a priori information about the state of the ocean is rather
  uncertain, there is a substantial efficiency gain in reacting to
  information obtained from measurements taken in-situ. %\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{Krause2008phd}.
\item For such types of sampling problems, the number of choices (i.e.
  locations, trajectories, and candidate designs) makes the problem
  combinatorially large, creating a trade-off between optimization
  (finding the most resource-efficient design to collect necessary
  data), and computability (arriving at a solution in reasonable
  time). To successfully resolve features, this trade-off has to be
  considered in sample design and practice.
\end{itemize}

@@

Central to understanding the changes taking place in the upper
water-column is an understanding of biophysical interaction driven by
an agglomeration of physical forcing (wind, topography, bathymetry,
tidal influences and local geography, such as fresh water inflows)
and incipient micro-biology driven by planktonic and coastal
anthropogenic input, such as pollution and agricultural runoff. These
often result in a range of ecosystem-related phenomena such as blooms,
anoxic zones and plumes, all of which have significant direct and indirect effects  on society. Walter Munk referred to the
20\textsuperscript{th} century as the \emph{century of undersampling}
\citep{munk2002}, pointing out a lack of sufficient sampling
resolution in time and space as a critical issue for characterizing
ocean phenomena. Even today, policymakers struggle with limited
observability of severe environmental problems such as the recent outbreak of harmful algal blooms in Northern
Norway\footnote{Report from the Institute of Marine
  Research: https://bit.ly/2KZGJfI} (May 2018), with losses exceeding 12 million
tons of salmon valued at \$80 million (Norwegian Directorate of
Fisheries). This is a case in point of how adequate sampling efforts
could have provided a more accurate assessment and early
warning of changing environmental conditions. These challenges related to undersampling have been driving
ocean sensing practices towards more autonomous, networked, and
distributed use of robotic assets in order to fill the gaps in our
understanding (See Fig. \ref{fig:envir}), where observation is
guided by assimilation of measurements and statistical models.

\begin{figure}[!h] 
  \centering 
  \subfigure[Illustration of an ocean sensing network, where robotic
  assets augment ship-based measurements to increase coverage and
  resolution of phenomena of scientific interest.]{\includegraphics[width =
    0.49\textwidth]{Figures/envir2.pdf}\label{fig:envir1}}
  \hfill
  \subfigure[Frontal patterns off of the Nidelva river, Trondheim, Norway.]{\includegraphics[width =
    0.49\textwidth]{Figures/river_proccess.pdf}\label{fig:nidelven}}
  \caption{Ocean observation is moving away from single-ship sampling
    towards more autonomous and collaborative networked operations in
    order to resolve the numerous bio-geochemical processes and their
    interactions.}
\label{fig:envir}
\end{figure}

@@

In this section, we calculate EPs and the Bernoulli variance for a few bivariate
Gaussian distributions to illustrate the concepts articulated above.
Fig. \ref{illus_bivarDens} shows contour plots of three different
densities with increasing correlation $\gamma$ between temperature and
salinity.
\begin{figure}[h!] \centering
  \includegraphics[width=0.99\textwidth]{Figures/illus_bivar.eps}
  \caption{Density contour plots with different correlations between
    temperature and salinity. The densities have unit variance and the
    thresholds are identical to the mean values $5^o C$ and
    $30 mg/l$. X-axis is temperature and y-axis is salinity.}
\label{illus_bivarDens}
\end{figure}
Here, the thresholds are set equal to the mean $\mu_T=t_T=5^o C$, and
salinity $\mu_S=t_S=30$ mg/l. The displayed densities have unit
standard deviations for both temperature and salinity, but we also
study the effect of doubling the standard deviations.

@@

\begin{figure}[!h] 
 \centering 
\includegraphics[width=0.98\textwidth]{Figures/field-trials/res_diag.pdf}
\caption{Data analysis from a preliminary trial experiment using the
  AUV. Left: Residual plot of temperature and salinity along with
  Gaussian contours. Middle: Empirical CDF (solid) of the quadratic form of
  the residuals along with the theoretical CDF (dashed) of the $\chi^2$
  distribution with two degrees of freedom. Right: Empirical variogram
  of the salinity and temperature data.} \label{fig:parest}
\end{figure}

Fig. \ref{fig:parest} shows diagnostic plots of this analysis. The display on the left shows a cross-plot of temperature and salinity residuals after the westernly trends in both salinity and temperature are subtracted from the data. This scatterplot of joint residuals indicates larger variability in salinity than in temperature, and a positive correlation ($0.5$) between the two variables. Based on the fitted bivariate Gaussian model (ellipses in left display), we can compute the modeled quadratic form of the residuals, and if the model is adequate they should be approximately $\chi^2_2$ distributed. Fig. \ref{fig:parest} (middle) shows the empirical cumulative distribution function (CDF) of these quadratic forms (solid) together with the theoretical CDF of the $\chi^2_2$ distribution. The modeled and theoretical curves are very similar, which indicates that the Gaussian model fits reasonably well. Even though there appears to be some clustering in both left and middle displays of Fig. \ref{fig:parest}, the bivariate diagnostic plots look reasonable and justify a Gaussian model. Fig. \ref{fig:parest} (right) shows the empirical variogram of the scaled residuals for temperature and salinity. The decay is similar for the two, and seems to be negligible after about $150$ m.

@@

\subfigure[Survey 1]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv1_es_sal.pdf}\label{fig:res1}}
\hspace{0.2cm}
\subfigure[Survey 2]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv4_es_sal.pdf}\label{fig:res2}}

\subfigure[ES for Survey 1]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_1.pdf}\label{fig:res3}}\hspace{0.4cm}
\subfigure[ES for Survey 2]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_4.pdf}\label{fig:res4}}
\caption{Results from mapping the Nidelva river} 
\label{fig:results}
\end{figure*}

@@

The final prediction of the front location given in Fig. \ref{fig:res3} and \ref{fig:res4}, correspond with one another, yielding a picture of the front gradually bending off toward the north east. The amount of exploration done by Survey 1 is greater than Survey 2. In Survey 1, the AUV obtained more detail by going north, while Survey 2, coming close to the survey area borders in the south-western corner, obtained a poorer understanding of the northern parts. A longer planning horizon would likely identify and discourage such choices.

@@

% == Adding references
\footnotesize
\bibliographystyle{apalike}
\bibliography{ref}

>>>>

% == Trygve Test
\usepackage{color, colortbl}
\definecolor{Gray}{gray}{0.9}

% put your definitions there:
\startlocaldefs
\endlocaldefs

@@

\begin{aug}
\author{\fnms{Trygve Olav} \snm{Fossum}\thanksref{t1,t2}, \corref{} \ead[label=e1]{trygve.o.fossum@ntnu.no}}
\author{\fnms{Jo} \snm{Eidsvik}\thanksref{t3}, \ead[label=e2]{jo.eidsvik@math.ntnu.no}}
\author{\fnms{David} \snm{Ginsbourger}\thanksref{t4}, \ead[label=e3]{david.ginsbourger@idiap.ch}}
\and
\author{\fnms{Kanna} \snm{Rajan}\thanksref{t2,t5,t6}. \ead[label=e4]{kanna.rajan@ntnu.no}}

\affiliation[t1]{Department of Marine Technology, The Norwegian University of Science and Technology (NTNU), Trondheim, Norway.} 
\affiliation[t2]{Centre for Autonomous Marine Operations and Systems (AMOS).}
\affiliation[t3]{Department of Mathematical Sciences, NTNU.}
\affiliation[t4]{Uncertainty Quantification and Optimal Design group, Idiap Research Institute, Martigny, Switzerland.}
\affiliation[t5]{Department of Engineering Cybernetics, NTNU}
\affiliation[t6]{Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}

\address{ \\Department of Marine Technology,\\ Trondheim, Norway \\
\printead{e1}}
\address{Department of Mathematical Sciences,\\ Trondheim, Norway\\ \printead{e2}}
\address{Idiap, \\ Martigny, Switzerland, \\
\printead{e3}}
\address{Department of Engineering Cybernetics,\\ Trondheim, Norway \\
\printead{e4}}

@@

The potential that lies in using information-based methods, which we highlight in the current paper, is to do efficient environmental sampling and overcome three common challenges, namely:

\begin{itemize}
\item Full numerical ocean models cannot provide accurate results online if run on robotic sensing platforms, as onboard computers cannot deliver the computational power required, hence statistical proxy models of the environment must be used for doing inference about where to sample.
\item As a priori information about the state of the ocean is uncertain, there is a substantial efficiency gain in reacting to information obtained from measurements taken in-situ. Recently acquired information can be assimilated into statistical models that is used to inform decisions on where to sample sequentially.%\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{Krause2008phd}.
\item For sampling problems related to environmental sensing, the number of choices (i.e. locations, trajectories, and candidate designs) makes the problem combinatorially large, creating a trade-off between optimization (finding the most resource-efficient design to collect necessary data), and computability (arriving at a solution in reasonable time). To successfully resolve features, this trade-off has to be considered in sample design and practice.
\end{itemize}

@@

Central to understanding the changes taking place in the upper
water-column is an understanding of biophysical interaction driven by
an agglomeration of physical forcing (wind, topography, bathymetry,
tidal influences and local geography, such as fresh water inflows)
and incipient micro-biology driven by planktonic and coastal
anthropogenic input, such as pollution and agricultural runoff. These
often result in a range of ecosystem-related phenomena such as blooms,
anoxic zones and plumes, all of which have significant direct and indirect effects  on society. Walter Munk referred to the
20\textsuperscript{th} century as the \emph{century of undersampling}
\citep{munk2002}, pointing out a lack of sufficient sampling
resolution in time and space as a critical issue for characterizing
ocean phenomena. Even today, policymakers struggle with limited
observability of severe environmental problems such as the recent outbreak of harmful algal blooms in Northern
Norway\footnote{Report from the Institute of Marine
  Research: https://bit.ly/2KZGJfI} (May 2018), with losses exceeding 12 million
tons of salmon valued at \$80 million (Norwegian Directorate of
Fisheries). This is a case in point of how adequate sampling efforts
could have provided a more accurate assessment and early
warning of changing environmental conditions. These challenges related to undersampling have been driving
ocean sensing practices towards more autonomous, networked, and
distributed use of robotic assets in order to fill the gaps in our
understanding (See Fig. \ref{fig:envir1}), where observation is
guided by assimilation of measurements and statistical models.

\begin{figure}[!h] 
  \centering 
  \subfigure[Illustration of an ocean sensing network and the sense-plan-act control methodology.]{\includegraphics[width =
    0.49\textwidth]{Figures/envir2.pdf}\label{fig:envir1}}
  \hfill
  \subfigure[Frontal patterns off of the Nidelva river, Trondheim, Norway.]{\includegraphics[width =
    0.49\textwidth]{Figures/river_proccess.pdf}\label{fig:nidelven}}
  \caption{\ref{fig:envir1} Ocean observation is moving away from single-ship sampling towards more autonomous and collaborative networked operations in order to resolve the numerous bio-geochemical processes and their interactions; this requires autonomy that can reason and make decisions on their own following the sense-plan-act control methodology. \ref{fig:nidelven} The interaction of river and ocean creates processes that are challenging to map, where the combination of robotics and statistics can play a vital role in enabling more effective observation practices.}
\label{fig:envir}
\end{figure}

@@

In this section, we calculate EPs and the Bernoulli variance for a few bivariate
Gaussian distributions to illustrate the concepts articulated above.
Fig. \ref{illus_bivarDens} shows contour plots of three different
densities with increasing correlation $\gamma$ between temperature and
salinity.
\begin{figure}[h!] \centering
  \includegraphics[width=0.99\textwidth]{Figures/illus_bivar.pdf}
  \caption{Density contour plots with different correlations between
    temperature and salinity. The densities have unit variance and the
    thresholds are identical to the mean values $5^o C$ and
    $30 mg/l$. X-axis is temperature and y-axis is salinity.}
\label{illus_bivarDens}
\end{figure}
Here, the thresholds are set equal to the mean $\mu_T=t_T=5^o C$, and
salinity $\mu_S=t_S=30$ mg/l. The displayed densities have unit
standard deviations for both temperature and salinity, but we also
study the effect of doubling the standard deviations.

@@

%\begin{figure}[!h] 
% \centering 
%\includegraphics[width=0.98\textwidth]{Figures/field-trials/res_diag.pdf}
%\caption{Data analysis from a preliminary trial experiment using the
%  AUV. Left: Residual plot of temperature and salinity along with
%  Gaussian contours. Middle: Empirical CDF (solid) of the quadratic form of
%  the residuals along with the theoretical CDF (dashed) of the $\chi^2$
%  distribution with two degrees of freedom. Right: Empirical variogram
%  of the salinity and temperature data.} \label{fig:parest}
%\end{figure}

\begin{figure}[!h]
  \centering
  \subfigure[Residual plot.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_a.pdf}\label{fig:parest_a}}
  \hfill
  \subfigure[Empirical CDF.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_b.pdf}\label{fig:parest_b}}
  \hfill
  \subfigure[Empirical variogram.]{\includegraphics[width = 0.32\textwidth]{Figures/field-trials/res_diag_c.pdf}\label{fig:parest_c}}
  \caption{Data analysis from a preliminary trial experiment using the AUV. \ref{fig:parest_a} Residual plot of temperature and salinity along with Gaussian contours. \ref{fig:parest_b} Empirical CDF (solid) of the quadratic form of the residuals along with the theoretical CDF (dashed) of the $\chi^2$  distribution with two degrees of freedom. \ref{fig:parest_c} Empirical variogram of the salinity and temperature data.} 
\label{fig:parest}
\end{figure}

Fig. \ref{fig:parest} shows diagnostic plots of this analysis. Fig. \ref{fig:parest_a} shows a cross-plot of temperature and salinity residuals after the westernly trends in both salinity and temperature are subtracted from the data. This scatterplot of joint residuals indicates larger variability in salinity than in temperature, and a positive correlation ($0.5$) between the two variables. Based on the fitted bivariate Gaussian model (ellipses in Fig. \ref{fig:parest_a}), we can compute the modeled quadratic form of the residuals, and if the model is adequate they should be approximately $\chi^2_2$ distributed. Fig. \ref{fig:parest_b} shows the empirical cumulative distribution function (CDF) of these quadratic forms (solid) together with the theoretical CDF of the $\chi^2_2$ distribution. The modeled and theoretical curves are very similar, which indicates that the Gaussian model fits reasonably well. Even though there appears to be some clustering in both Fig. \ref{fig:parest_a} and \ref{fig:parest_b}, the bivariate diagnostic plots look reasonable and justify a Gaussian model. Fig. \ref{fig:parest_c} shows the empirical variogram of the scaled residuals for temperature and salinity. The decay is similar for the two, and seems to be negligible after about $150$ m.

@@

\subfigure[Survey 1]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv1_es_sal_ep.pdf}\label{fig:res1}}
\hspace{0.2cm}
\subfigure[Survey 2]{\includegraphics[height=0.40\textwidth]{Figures/field-trials/auv4_es_sal_ep.pdf}\label{fig:res2}}

%\subfigure[ES for Survey 1]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_1.pdf}\label{fig:res3}}\hspace{0.4cm}
%\subfigure[ES for Survey 2]{\includegraphics[height=0.41\textwidth]{Figures/field-trials/ep_4.pdf}\label{fig:res4}}
\caption{Results from mapping the Nidelva river. \ref{fig:map} Show an overview of the survey area overlaid with the AUV path in black and dashed line. Note the shaded region indicating the usual frontal region. \ref{fig:res_both} Presents the collected temperature data as colored trails. Note waypoint 5 (WP5) which indicate where the two surveys deviate. \ref{fig:res1} and \ref{fig:res2} Shows the collected salinity data (as a colored trail) overlaid on the final EP, which indicate the AUVs statistical impression of the front. For both missions the temperature and salinity data correspond with the EP front indication.} 
\label{fig:results}
\end{figure*}

@@

The final prediction of the front location given in Fig. \ref{fig:res1} and \ref{fig:res2} as a dashed line, correspond with one another, yielding a picture of the front gradually bending off toward the north east. The amount of exploration done by Survey 1 is greater than Survey 2. In Survey 1, the AUV obtained more detail by going north, while Survey 2, coming close to the survey area borders in the south-western corner, obtained a poorer understanding of the northern parts. A longer planning horizon would likely identify and discourage such choices.

@@

% == Adding references
\footnotesize
\bibliographystyle{imsart-nameyear}
\bibliography{ref}
